Determinism, Complexity, and Predictability in Computer Performance

### 2014-08-04 response.tex

We thank the reviewer for their kind words and appreciate the time
they put into reviewing our work.

\noindent\emph{Major issues:}

\emph{I don't get the reason for normalizing the three methods by the
  random walk method (for example in Table I). As the authors write
  themselves this makes the indicator vulnerable to a bias caused by
  the high influence of this (arbitrarily selected) reference
  method. Why not just use a non-referential estimator, which then
  would also allow to compare the inherent predictability of different
  systems (as it is now the values obtained for each systems are
  strongly influenced by the random walk estimate on that particular
  system and thus can not be compared). This will also change the
  results of the fitting (the authors explicitly exclude one value
  because of the effect caused by the normalization). And it is not
  that the random walk predictor is the one standard method that all
  other methods have to compete against.}

No error metric is perfect.  For our purposes, we needed one that was
scale free and comparable across different time series.  While MASE
has its drawbacks, it was a very good fit to those requirements.  In
our opinion, this outweighed its disadvantages.  There is a great deal
of discussion of this cluster of issues in [Hyndman and Koehler, 2006]
(our reference [40]).  If the reviewer thinks that it would strengthen
our paper, we can recapitulate some of that material in our Section
IVD.

@@

This short paragraph in this paper, which explains the efficiencies
and inefficiencies of these schemes, was intended to serve two
purposes: first, to explain why we have included these simple
prediction schemes and not just the more-complicated ones; second, to
help in explaining the specific results described later in the paper.

\smallskip

\emph{9) A better justification of the prediction error metrics could
  be also in order if this work is continued. At least, comparison of
  prediction errors quantified with different metrics seems
  necessary. Especially, taking into account the following major
  concerns.}

We have extended the justification for our use of the MASE metric; see
the last paragraph in Section VI.  As mentioned there, the paper in
which MASE was introduced---our reference [40]---includes an extensive
evaluation of this metric, as well as a comprehensive comparison to
other metrics.  If the reviewer thinks that it would strengthen our
paper, we can recapitulate more of that material in our paper.  As
part of our current work on this topic, we are evaluating how other
choices of metric affect the shape of our curves, but we feel that a
comprehensive study of this issue is best left as a sequel to this
work.

@@

The WPE vs. MASE points of the transformed data move---upwards,
generally---but they remain within the shaded region.  Indeed, the
values for the transformed version of {\tt dgesdd}$_5$ (dark-green
icons) move {\sl closer} to the curve.

This series of experiments supports our claims of generality: a
transformation like this effectively changes the generating process,
but does not affect the performance of our model-free heuristic.  And
the curves that capture that heuristic---which were derived from data
from a single computer-performance experiment---do not only hold for
traces from that experiment and transformed versions of them.  The
preliminary results of our broader study, currently underway, indicate
that they also hold for data from synthetic examples and for
real-world data from a different system.  The image below shows some
of these results:

\begin{center}
    \includegraphics[width=0.8\columnwidth]{figs/new_prediction_vs_entropy_extras_with_nonlinear_all_points}
\end{center}

The blue fit lines here are derived from the original 120 traces; the
magenta one is fit to those traces plus the nonlinear transformations
of {\tt dgesdd}$_2$, {\tt dgesdd}$_5$, and {\tt dgesdd}$_6$, the $x$
coordinate of a 10,000-point orbit from the H\'{e}non map with $a=1.4$
and $b=0.3$, the $x$ coordinate of a 10,000-step 4th-order Runge-Kutta
solution of the Lorenz equations with $r=28, \; \sigma=10, \; b=8/3$
and $\Delta t=0.01$, a 10,000-step random-walk time series generated
by taking the cumulative sum of $\mu=0$, $\sigma=1$ Gaussian random
variables using Python's {\tt numpy} library\footnote{{\tt
    numpy.cumsum(numpy.random.normal(0.0, 1.0, size=10000))}}, and
  data set ``A'' from the Santa Fe Institute time-series competition
  website.

### 2014-08-04 concl.tex
conclusion hacking

@@

In this paper, we have argued that redundancy~\cite{crutchfield2003}
is a useful definition of the inherent predictability of an empirical
time series.  To operationalize that definition, we use an
approximation of the Kolmogorov-Sinai entropy~\cite{lind95}, estimated
using a weighted version of the permutation entropy
of~\cite{bandt2002per}.  This WPE technique---an ordinal calculation
of forward information transfer in a time series---is ideal for our
purposes because it works with real-valued data and is known to
converge to the true entropy value. Using a variety of forecast models
and more than 150 time-series data sets from experiments and
simulations, we have shown that prediction accuracy is indeed
correlated with weighted permutation entropy: the higher the WPE, in
general, the higher the prediction error.  The relationship is roughly
logarithmic, which makes theoretical sense, given the nature of WPE,
predictability, and MASE.

@@

These curves were determined empirically using a specific error metric
and a finite set of forecast methods and time series traces.  If one
uses a different error metric, the geometry---and the heuristic---will
no longer apply.  And while the methods and traces were chosen to be
representative of the practice, they are of course not completely
comprehensive.  It is certainly possible, for instance, that the
nonlinear dynamics of computer performance is subtly different from
the nonlinear dynamics of other systems.  Our preliminary results on
other systems (H\'enon, Lorenz, white noise, SFI ``A'', transformations
of the computer performance data) lead us to believe that our results
will generalize beyond the examples described in this paper.  We are
in the process of following up on that exploration with a broader
study of data, forecast methods, and error metrics.

@@

Nonstationarity is a serious challenge in any time-series modeling
problem.  [[Can destroy our heuristic: regime shifts can cause the
    test signal to look different from the training signal, which can
    skew the MASE score (and the points on the curve).  A RW MASE
    score larger than 1, however, as mentioned at end of prev section,
    can flag when this is happening.]]  Detecting regime shifts---and
adapting prediction models accordingly---is an important area of
future work.  Indeed, one of the first applications of permutation
entropy was to recognize the regime shift in brainwave data that
occurs when someone has a seizure~\cite{cao2004det}.  Recall that the
signal in Figure~\ref{fig:svd-ts-colored} was especially useful for
the study in this paper because it contained a number of different
regimes.  We segmented this signal visually, but one could imagine
using permutation entropy to do so instead.  Automating regime-shift
detection would be an important step towards a fully adaptive modeling
strategy, where old models are discarded and new ones are rebuilt
whenever the time series enters a new regime.  WPE would be
particularly powerful in this scenario, as its value can not only help
with regime-shift detection, but also suggest what kind of model might
work well in each new regime.  Of particular interest would be the
class of so-called \emph{hybrid systems}~\cite{hybrid}, which exhibit
discrete transitions between different continuous regimes---e.g., a
lathe that has an intermittent instability or traffic at an internet
router, whose characteristic normal traffic patterns shift radically
during an attack.  Effective modeling and prediction of these kinds of
systems is quite difficult; doing so adaptively and automatically---in
the manner that is alluded to at the end of the previous
paragraph---would be an even more important challenge.

>>>>

\alert{add FW stuff from section VI}

@@

In this paper, we have argued that redundancy is a useful definition
of the inherent predictability of an empirical time series.  To
operationalize that definition, we use an approximation of the
Kolmogorov-Sinai entropy, estimated using a weighted version of the
permutation entropy of~\cite{bandt2002per}.  This WPE technique---an
ordinal calculation of forward information transfer in a time
series---is ideal for our purposes because it works with real-valued
data and is known to converge to the true entropy value. Using a
variety of forecast models and more than 150 time-series data sets
from experiments and simulations, we have shown that prediction
accuracy is indeed correlated with weighted permutation entropy: the
higher the WPE, in general, the higher the prediction error.  The
relationship is roughly logarithmic, which makes theoretical sense,
given the nature of WPE, predictability, and MASE.

@@

These curves were determined empirically using a specific error metric
and a finite set of forecast methods and time series traces.  If one
uses a different error metric, the geometry---and the heuristic---will
no longer apply.  And while the methods and traces were chosen to be
representative of the practice, they are of course not completely
comprehensive.  It is certainly possible, for instance, that the
nonlinear dynamics of computer performance is subtly different from
the nonlinear dynamics of other systems.  Our preliminary results on
other systems (H\'enon, Lorenz, a white-noise process, SFI ``A'',
transformations of the computer performance data) lead us to believe
that our results will generalize beyond the examples described in this
paper.  We are in the process of following up on that exploration with
a broader study of data, forecast methods, and error metrics.

@@

Nonstationarity is a serious challenge in any time-series modeling
problem.  Any regime shift that causes a change in the predictive
structure between the training signal and the test signal, for
instance, will skew the MASE score (and the points on the curve).
This can destroy the utility of our heuristic.  \alert{Conversely,
  though,} a MASE score of a random-walk prediction that is larger
than 1---as mentioned at end of the previous section---can potentially
be a good indicator that a signal is nonstationary.  Detecting regime
shifts---and adapting prediction models accordingly---is an important
area of future work.  Indeed, one of the first applications of
permutation entropy was to recognize the regime shift in brainwave
data that occurs when someone has a seizure~\cite{cao2004det}.  Recall
that the signal in Figure~\ref{fig:svd-ts-colored} was especially
useful for the study in this paper because it contained a number of
different regimes.  We segmented this signal visually, but one could
imagine using permutation entropy to do so instead.  Automating
regime-shift detection would be an important step towards a fully
adaptive modeling strategy, where old models are discarded and new
ones are rebuilt whenever the time series enters a new regime.  WPE
would be particularly powerful in this scenario, as its value can not
only help with regime-shift detection, but also suggest what kind of
model might work well in each new regime.  Of particular interest
would be the class of so-called \emph{hybrid systems}~\cite{hybrid},
which exhibit discrete transitions between different continuous
regimes---e.g., a lathe that has an intermittent instability or
traffic at an internet router, whose characteristic normal traffic
patterns shift radically during an attack.  Effective modeling and
prediction of these kinds of systems is quite difficult; doing so
adaptively and automatically---in the manner that is alluded to at the
end of the previous paragraph---would be an interesting challenge.
